Paper template for {ICPhS} 2023 Prague

### 2024-03-22 manuscript/main.Rmd

---
documentclass: article
classoption: [a4paper,11pt,twocolumn]
subparagraph: yes
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1]{fontenc}
  #- \usepackage[leqno,fleqn]{amsmath}
output:
  pdf_document:
    highlight: zenburn
    includes:
      before_body: includes/tex/doc_prefix.tex
      in_header: includes/tex/header.tex
    keep_tex: yes
    number_sections: yes
bibliography: includes/bib/icphs2023.bib
---

@@

\begin{itemize}
  \item they must be written in English
  \item they must not contain the name(s) of the author(s) (for
anonymous review)
  \item the maximum is four pages for Congress papers, plus up to one additional page of references. Papers for Plenary lectures can be longer
  \item they must be submitted in PDF format (cf.\ Section~4)
  \item the paper submission will occur via a web interface
\end{itemize}
This paper template can be found on the conference website. If there
are special questions or wishes regarding paper preparation and
submission for ICPhS 2023, correspondence should be addressed to
icphs2023@guarant.cz. Information for full paper submission will be
available on the web at \url{https://www.icphs2023.org/}.

@@

Write something?

@@

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015). In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p. 279). In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [sɛl] - *zel* [zɛl] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019). Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022). For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tórəs] and *torres antigues* [torəzəntiɣəs].

@@

>>>>

---
documentclass: article
classoption: [a4paper,11pt,twocolumn]
subparagraph: yes
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage[T1]{fontenc}
  - \usepackage{tipa}
  #- \usepackage[leqno,fleqn]{amsmath}
output:
  pdf_document:
    highlight: zenburn
    includes:
      before_body: includes/tex/doc_prefix.tex
      in_header: includes/tex/header.tex
    keep_tex: yes
    number_sections: yes
bibliography: includes/bib/icphs2023.bib
---

@@

@@

@@

Both Catalan and Peninsular Spanish sound inventories have an phoneme /s/, which can be classified as a voiceless alveolar fricative (Hualde & Colina, 2015). In addition, most of the Catalan varieties have a voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens Vives, 1996, p. 279). In onset position, /s/ and /z/ contrast phonemically, which leads to minimal pairs like *cel* [s\textepsilon l] - *zel* [z\textepsilon l] and *fossa* [fósa] - *fosa* [fóza] (Carrera-Sabaté & Prieto, 2012-2019). Conversely, at the end of the word, this contrast is neutralized, being the voicing or devoicing of the alveolar fricative dependent on the following sound (Institut d’Estudis Catalans, 2022). For example, if it is followed by a vowel, the alveolar fricative is voiced: *torres* [tór\textschwa{}s] and *torres antigues* [tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s].

@@

\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}

The social meaning and the perception of intervocalic [z] in Spanish has been studied in a variety of contexts, including both monolingual (Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell & García, 2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts (Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et al., 2023; García, 2015)With respect to the former, it has been found that for listeners from San José, Costa Rica, the covert attitudes associated with intervocalic [z] vary significantly by speaker gender. Consequently, male speakers who use [z] tend to be positively evaluated, and generally perceived as nice, local, confident, and masculine. Conversely, female speakers who use [z] are only afforded negative associations of lower education and social class (Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).

Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is evaluated differently in male and female speakers. Accordingly, women who produce intervocalic [z] are viewed as significantly lower status, less pleasant, and older. Contrary, men who produce intervocalic [z] do not receive these negative evaluations. Additionally, this variable has been found to serve as a marker of regional identity, with voicing being associated with neighboring Highland Ecuadorian dialects (Cuenca and Quito) that exhibit systematic voicing. García (2020) hypothesizes that young men are the leaders of this change, essentially due to the fact that they are in greater contact with these highland varieties. Interestingly, the mean voicing rates of females are roughly two age groups behind those of males. Consequently, it is believed that it is not necessarily the case that women in Ecuador are avoiding voicing like San José women due to its negative evaluations, but rather that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic [z] in bilingual contexts, this phenomenon has been mainly studied in Catalan-Spanish contact situations, given its robust usage by Catalan-Spanish bilinguals and, in particular, by those with greatest exposure to and usage of Catalan (Davidson, 2014, 2015). According to Sinner (2002), intervocalic [z] can be considered a possible linguistic marker of Catalonian Spanish, despite its relative lack of social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner (2002) interviewed 12 speakers of Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of linguistic features of Catalanized Spanish. Results showed that only select Barcelona speakers perceived intervocalic [z] as a linguistic marker of Catalonian Spanish, whereas monolingual Spanish-speakers from Madrid did not. Additionally, Davidson (2019) found that, in Barcelona, the indexical field of intervocalic [z] features exclusively positive associations with bilingualism and local solidarity, which warrants its gradual adoption in the speech community, unlike other phonetic markers of Catalonian Spanish that are the topic of overt social ridicule, such as lateral velarization. These results are in line with previous findings on intervocalic [z] in Catalonian Spanish. For instance, Silverstein (2003) found that intervocalic [z] indexes a local Barcelona identity, and the variant then acquires positive associations with bilingualism and cultural pride. Overall, these findings suggest that intervocalic [z] likely functions as a sociolinguistic indicator of Catalonian Spanish within the Catalonian speech community, indexing Catalan-Spanish bilingual membership while lacking any considerable degree of generalized conscious awareness (Davidson, 2019).

### 2024-03-22 manuscript/main.tex

% Packages removed from icphs2023.sty because of conflicts
% They have been added to the .Rmd yaml front matter
% \usepackage[latin1]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage[leqno,fleqn]{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{itemize}
  \item they must be written in English
  \item they must not contain the name(s) of the author(s) (for
anonymous review)
  \item the maximum is four pages for Congress papers, plus up to one additional page of references. Papers for Plenary lectures can be longer
  \item they must be submitted in PDF format (cf.\ Section~4)
  \item the paper submission will occur via a web interface
\end{itemize}

This paper template can be found on the conference website. If there are
special questions or wishes regarding paper preparation and submission
for ICPhS 2023, correspondence should be addressed to
\href{mailto:icphs2023@guarant.cz}{\nolinkurl{icphs2023@guarant.cz}}.
Information for full paper submission will be available on the web at
\url{https://www.icphs2023.org/}.

@@

Both Catalan and Peninsular Spanish sound inventories have an phoneme
/s/, which can be classified as a voiceless alveolar fricative (Hualde
\& Colina, 2015). In addition, most of the Catalan varieties have a
voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens
Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
phonemically, which leads to minimal pairs like \emph{cel} {[}sɛl{]} -
\emph{zel} {[}zɛl{]} and \emph{fossa} {[}fósa{]} - \emph{fosa}
{[}fóza{]} (Carrera-Sabaté \& Prieto, 2012-2019). Conversely, at the end
of the word, this contrast is neutralized, being the voicing or
devoicing of the alveolar fricative dependent on the following sound
(Institut d'Estudis Catalans, 2022). For example, if it is followed by a
vowel, the alveolar fricative is voiced: \emph{torres} {[}tórəs{]} and
\emph{torres antigues} {[}torəzəntiɣəs{]}.

@@

>>>>

% Packages removed from icphs2023.sty because of conflicts
% They have been added to the .Rmd yaml front matter
% \usepackage[latin1]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage[leqno,fleqn]{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tipa}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

@@

Both Catalan and Peninsular Spanish sound inventories have an phoneme
/s/, which can be classified as a voiceless alveolar fricative (Hualde
\& Colina, 2015). In addition, most of the Catalan varieties have a
voiced alveolar phoneme in intervocalic contexts (Lloret, 2011; Recasens
Vives, 1996, p.~279). In onset position, /s/ and /z/ contrast
phonemically, which leads to minimal pairs like \emph{cel}
{[}s\textepsilon l{]} - \emph{zel} {[}z\textepsilon l{]} and
\emph{fossa} {[}fósa{]} - \emph{fosa} {[}fóza{]} (Carrera-Sabaté \&
Prieto, 2012-2019). Conversely, at the end of the word, this contrast is
neutralized, being the voicing or devoicing of the alveolar fricative
dependent on the following sound (Institut d'Estudis Catalans, 2022).
For example, if it is followed by a vowel, the alveolar fricative is
voiced: \emph{torres} {[}tór\textschwa{}s{]} and \emph{torres antigues}
{[}tor\textschwa{}z\textschwa{}nti\textgamma{}\textschwa{}s{]}.

@@

\subsection{Language attitudes towards the alveolar fricative in Catalan and Spanish}

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \& García,
2017; García, 2015, 2019; Silverstein, 2003) and bilingual contexts
(Davidson, 2014, 2015, 2019; Sinner, 2002), despite having been
categorized as a low-salient variant (Chappell, 2017, 2021; Chappell et
al., 2023; García, 2015)With respect to the former, it has been found
that for listeners from San José, Costa Rica, the covert attitudes
associated with intervocalic {[}z{]} vary significantly by speaker
gender. Consequently, male speakers who use {[}z{]} tend to be
positively evaluated, and generally perceived as nice, local, confident,
and masculine. Conversely, female speakers who use {[}z{]} are only
afforded negative associations of lower education and social class
(Chappell, 2016; Chappell et al., 2023; Silverstein, 2003).

Similarly, García (2019) shows that, in Ecuador, intervocalic /s/ is
evaluated differently in male and female speakers. Accordingly, women
who produce intervocalic {[}z{]} are viewed as significantly lower
status, less pleasant, and older. Contrary, men who produce intervocalic
{[}z{]} do not receive these negative evaluations. Additionally, this
variable has been found to serve as a marker of regional identity, with
voicing being associated with neighboring Highland Ecuadorian dialects
(Cuenca and Quito) that exhibit systematic voicing. García (2020)
hypothesizes that young men are the leaders of this change, essentially
due to the fact that they are in greater contact with these highland
varieties. Interestingly, the mean voicing rates of females are roughly
two age groups behind those of males. Consequently, it is believed that
it is not necessarily the case that women in Ecuador are avoiding
voicing like San José women due to its negative evaluations, but rather
that they are further behind in the adoption of the incoming variant.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations, given its robust usage by
Catalan-Spanish bilinguals and, in particular, by those with greatest
exposure to and usage of Catalan (Davidson, 2014, 2015). According to
Sinner (2002), intervocalic {[}z{]} can be considered a possible
linguistic marker of Catalonian Spanish, despite its relative lack of
social awareness (Davidson, 2019; Sinner, 2002). In their study, Sinner
(2002) interviewed 12 speakers of Barcelonan Spanish and monolingual
Spanish-speakers from Madrid aged 27 to 41 regarding their awareness of
linguistic features of Catalanized Spanish. Results showed that only
select Barcelona speakers perceived intervocalic {[}z{]} as a linguistic
marker of Catalonian Spanish, whereas monolingual Spanish-speakers from
Madrid did not. Additionally, Davidson (2019) found that, in Barcelona,
the indexical field of intervocalic {[}z{]} features exclusively
positive associations with bilingualism and local solidarity, which
warrants its gradual adoption in the speech community, unlike other
phonetic markers of Catalonian Spanish that are the topic of overt
social ridicule, such as lateral velarization. These results are in line
with previous findings on intervocalic {[}z{]} in Catalonian Spanish.
For instance, Silverstein (2003) found that intervocalic {[}z{]} indexes
a local Barcelona identity, and the variant then acquires positive
associations with bilingualism and cultural pride. Overall, these
findings suggest that intervocalic {[}z{]} likely functions as a
sociolinguistic indicator of Catalonian Spanish within the Catalonian
speech community, indexing Catalan-Spanish bilingual membership while
lacking any considerable degree of generalized conscious awareness
(Davidson, 2019).
